# Publishing Your Dataset {#sec-datapublication}

```{r, include=FALSE}
library(washr)
```

## Creating and Editing the README

To start documenting your dataset for publication, you first need to set up a README file.

### Initialize the README

In your R console, within the project directory, run the following command to create a README file:

```{r}
setup_readme()
```


If you want the README to include an Example section with a scaffold for a first plot of your data, use this command instead:

```{r}
setup_readme(has_example = TRUE)
```

The template documents the first data object in `data/`. When your package holds several, copy the data section once per further object.

### Editing the README

Locate the `README.Rmd` file in your project directory and open it. Edit each section of the README to provide relevant information about your package. Typical sections include:

- **Package Name and Brief Description**
- **Installation Instructions**
- **Basic Usage**
- **Features**
- **Example**

### Creating a Data Visualization

In the "Example" section of your README, it’s helpful to include at least one plot that showcases your data. The section written by `setup_readme(has_example = TRUE)` holds a commented scaffold for it. For instance, using `ggplot2`, you could create a plot like this (adapt it to fit your specific data):

```{r}
library(ggplot2)
library(yourpackagename)

ggplot(your_data, aes(x = variable1, y = variable2)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Example Plot from YourPackageName")
```

### Building the README

Once you’ve completed editing the README, convert it to a Markdown file by running:

```{r}
devtools::build_readme()
```

This will generate a `README.md` file, which GitHub displays on your repository's main page.

### Updating GitHub

After building the README, go to the "Git" tab in RStudio. Stage the newly created or modified README files, commit them with a message such as "Add and update README," and push the changes to GitHub.

## Setting Up the Package Website

To enhance your package's accessibility, you can create a website for it.

### Initializing the Website

Run the following command in your R console:

```{r}
setup_website()
```

The first run writes `_pkgdown.yml` from the openwashdata template and builds the site into `docs/`. Every later run keeps your `_pkgdown.yml` as it is and rebuilds the site, so there is nothing to answer.

### Applying the openwashdata brand

To give the site the openwashdata fonts and colors, run:

```{r}
use_brand()
```

This copies the brand file and the logos from the central brand repository into your package and points `_pkgdown.yml` at them. Run it again whenever the brand changes.

### Document, Check, and Install the Package

Ensure your package is up-to-date by running the following commands:

```{r}
devtools::document()
devtools::check()
devtools::install()
```

These commands generate the latest documentation, check the package for issues, and install it locally.

### Preparing for GitHub Pages

To host the package website on GitHub Pages, Git has to track the built site. `setup_website()` removes the `docs` line from `.gitignore` for you, so the `docs/` folder is tracked. Check that the line is gone before you commit.

### Updating GitHub with Website Files

Go to the "Git" tab in RStudio. You should now see new files in the `docs/` folder and the updated `.gitignore` file. Stage these changes, commit them with a message like "Add pkgdown website files," and push the updates to GitHub.

### Setting Up GitHub Pages

To activate the website, open your GitHub repository in a web browser. Go to "Settings" > "Pages." Under "Source," select the branch where your `docs/` folder resides (usually "main" or "master") and set the folder to `/docs`. Finally, click "Save."

Your package website will now be live at `https://yourusername.github.io/yourrepositoryname/`.

Remember, if you make significant changes to your package or documentation, you may need to rebuild the website by running `pkgdown::build_site()` and pushing the updated `docs/` folder to GitHub.